---
title: "data-viz"
author: "Scout Leonard"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#General use packages
library(tidyverse)
library(here)
library(janitor)

#Specifically for data visualization
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(sf)
```

## Get the lizards data

```{r}
lizards <- read_csv(here::here("data_tidy", "lizards.csv"))
```

## Make a histogram 

Make a histogram of lizard weights.

If we want to fill in the graph:
- **color** refers to points and lines
- **fill** refers to the color of polygons
- **size** changes the thickness of your lines 
- **linetype** defines the visualization of your lines 

```{r}
ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(fill = "orange",
                 color = "cyan",
                 size = 0.5,
                 linetype = "dotted")
```
## Point Plots

Make a point plot (scatterplot) with total length on x axis and weight on y axis. Shape (0-25)

*Different shape numbers, from 0 to 25, represent different point shapes.*

- **alpha** gives the opacity

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(shape = 21,
             fill = "yellow",
             color = "red",
             size = 4,
             alpha = 0.5)
```

## Bubble plot 

Make a bubble plot where the color of the points varies based on common_name, and the size of points varies based in total_length.

DISCLAIMER: BAD IDEA

*when something changes based on a variable, it goes inside `aes()`. If something changes based on a constant, it goes outside `aes()`. 

**There are a number of themes you can include in your graphs** 

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, size = total_length),
             shape = 22) +
  theme_dark()
```

## Facet wraps 

Helps people to see trends within groups. Groups the desired variable into groups. 

You can choose the number of columns to help visualization. 

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_wrap(~common_name, ncol = 4)
```
## Facet Grids

Creates grids based on the intersection of variables. 

*Different levels of the sex variables in rows and tail variables in columns*

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name)) +
  facet_grid(sex ~ tail)
  theme_light()
```
If you have a character, ggplot automattically arranges them alphabetically. 

Find total lizard counts by different groups in common name. 
```{r}
lizards_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(count = n())

lizards_counts

#Same Thing 
# lizards_counts <- lizards %>% dplyr::count(common_name)

#You can also use it to get the counts of two variables! 
# lizard_counts_cn_tail <- lizards %>%  dplyr::count(common_name, tail)

```

To plot the total counts for lizards in this dataset: 

*Put long categorical names in the y axis for space!* 

Or use `coord_flip()`

The common names show up in alphabetical order. We can make common name a factor(instead of character), and assign the factor levels so they show up in a certain order. `fct_reorder()` makes the variable a factor. The "count" specification tells ggplot to reorder them by count in each variable! 

```{r}
ggplot(data = lizards_counts, aes(y = fct_reorder(common_name, count), x = count)) +
  geom_col(aes(fill = common_name), show.legend = FALSE) +
  labs(x = "Lizard Counts",
       )
```
Try converting common name to an ordered factor outside of ggplot, so that the factors are stored. 

```{r}
lizards_counts <- lizards_counts %>% 
  mutate(common_name = fct_reorder(common_name, count))
         
class(lizards_counts$common_name)


#SHows that the levels are no longer alphabetical
levels(lizards_counts$common_name)
```
## Axis scale breaks, limits, and labels 

Make a scatterplot of total_length versus weight. 

- **Continuous variables**
  - breaks are the tick mark breaks 
  - expands beyond your limits! but you can set expand to zero to eliminate that 
    - When you see the outcome, you can see that removing the exmpansion can sometimes cut off points: sometimes it's valuable! 
```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(0, 10, 50, 500),
                     limits = c(0, 500),
                     expand = c(0,0))+
  scale_y_continuous(expand = c(0,0))

```

Create a new version of lizards where lizards are counted by date.
Transform date column to class Date, then find counts of observation by date.
```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(date)
  
```

- **Date**
Make a plot now that date has been converted to class date. 
Make a line plot (geom_line) of date (x) and count(y)
Know what axes you are working on!
Know what kind of data you are working with! 
```{r}
ggplot(data = lizard_counts, aes(x = date, y = n)) +
  geom_line() + 
  scale_x_date(date_breaks = "3 years",
               date_labels = ("%y"))
```

- **Discrete values**





